Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS

### 2025-05-16 report/Chenning_paper.qmd
update

---
title: "Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS"
format: 
    pdf:
        documentclass: article 
        number-sections: true
        margin-top: 1in
        margin-bottom: 1in
        margin-left: 1in
        margin-right: 1in
        linestretch:  1.2
        fontsize: 10pt
        fig-pos: "H"
        author: "Chenning Xu"
bibliography: tom.bib
---

@@

China's economic trajectory over the past five decades, initiated by the reform and opening-up policy launched in 1978, involved a transition from a centrally planned to a more market-oriented economy. This period yielded high average GDP growth rates averaged over 9% per year along with considerable improvements in poverty reduction, health, and education [@world_bank_world_2024]. This growth, however, manifested unevenly across time, geography, and economic sectors. Coastal provinces, for instance, often experienced higher growth rates, benefiting from earlier integration with global markets, while inland and rural areas developed more slowly, contributing to regional disparities [@liao_sixty_2016]. Sectorally, growth was initially propelled by fixed-capital investment and export-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024]. China's annual GDP growth rate moderated from 10.6% in 2010 to 5.2% by 2023, reflecting a slowdown from the high rates experienced earlier in the decade [@world_bank_gdp_2025].

@@

This study contributes to the understanding of China's wealth inequality by employing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS). The analysis focuses specifically on assessing whether wealth inequality trends during this later period substantiate claims of ongoing inequality reduction observed in earlier research focused primarily on income [@kanbur_great_2021; @zhang_survey_2021]. The paper utilizes factor decomposition methods to pinpoint the contributions of various asset types to overall wealth inequality. A further level of analysis involves applying this factor decomposition separately to distinct population subgroups defined by urban-rural residence and coastal-inland location, allowing for an examination of potential heterogeneity in the drivers of inequality across these dimensions. Additionally, group decomposition techniques are used to quantify the portion of total wealth inequality attributable to the average wealth differences between these urban-rural and coastal-inland groups. This combination of recent data and multiple decomposition strategies aims to offer a detailed perspective on the patterns and structural underpinnings of wealth inequality in contemporary China.

@@

The CHFS provides sample weights, calculated based on the multi-stage sampling design and adjusted using population benchmarks, to facilitate representative analysis. The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior. Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others.

@@

## Factor decomposition of wealth inequality
This section employs the decomposition method developed by @lerman_income_1985 to identify the specific contributions of different asset types to overall wealth inequality in China. The decomposition facilitates a detailed examination of how various asset components individually contribute to the aggregate wealth inequality measured by the Gini coefficient. This approach enables an analysis of the structural composition of inequality, clarifying not only the relative magnitude of each asset type's impact but also the extent to which each correlates with the overall wealth distribution. According to this method, the overall wealth Gini coefficient can be decomposed into contributions from each asset component as follows:

@@

where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth. This decomposition clarifies not only how much each asset type contributes to total inequality based on its relative size and internal inequality, but also how strongly each asset correlates with the overall wealth ranking. The summary statistics of each asset type are presented in @tbl-wealth-summary.

@@

Relative to the nationwide decompositions for 2017 and 2019, the regional analyses reaffirm the pervasive significance of housing, land, and financial assets in shaping wealth inequality across China, while simultaneously highlighting distinctive regional patterns. In the coastal East, the influence of housing is particularly pronounced, contributing over 80 percent (80.3%) of total inequality—approximately eight percentage points above the national average. In contrast, housing accounts for a comparatively smaller share (58.3%) in the Non-East regions. Across all three decomposition components, housing distribution in the East aligns more closely with overall wealth rankings, is more unequal, and constitutes a greater share of total household wealth.

@@

The above summary statistics shows that the mean and median wealth in the East is significantly higher than in the other regions in both 2017 and 2019. Notably, the mean wealth of West-area households surpassed that of the Middle region in 2019, although its median wealth remained lower. Concurrently, the West's Gini coefficient rose to become the highest among all regions in 2019 (0.6947).

@@

The group decomposition indicates that, analogous to the urban-rural analysis, the between-group component accounted for over 30% of total wealth inequality in both 2017 and 2019 (36.9% and 33.8%, respectively). Between-group contributions experienced a slight decline from 2017 to 2019, mirrored by a corresponding increase in the residual component (from 32.0% to 34.5%). This shift may suggest a modest convergence in household wealth across regions. Nonetheless, it remains clear that China's wealth inequality arises significantly from both within-region disparities and regional differences, with each component consistently contributing over 30% to overall inequality.

# Conclusion and policy implications
This paper analyzed wealth inequality in China using data from the 2017 and 2019 China Household Finance Survey (CHFS). The findings confirm arguments from earlier studies on China's income inequality, indicating that the country's wealth inequality has plateaued since 2010. Specifically, the national wealth Gini coefficient declined from 0.76 in 2010 to 0.70 in 2017, and further decreased to 0.67 in 2019.

>>>>

---
title: "Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS"
format: 
    pdf:
        documentclass: article 
        number-sections: true
        margin-top: 1in
        margin-bottom: 1in
        margin-left: 1in
        margin-right: 1in
        linestretch:  1.2
        fontsize: 12pt
        author: "Chenning Xu"
        fig-pos: "H"
bibliography: tom.bib
abstract: "This paper analyzes recent trends and structural determinants of wealth inequality in China using microdata from the 2017 and 2019 China Household Finance Survey (CHFS). The findings indicate a moderation in wealth inequality, aligning with earlier research focusing primarily on income inequality. By employing factor decomposition and group decomposition techniques, the analysis highlights considerable heterogeneity in the profiles of wealth inequality between urban-rural and coastal-inland populations. It further demonstrates that both urban-rural and regional disparities substantially contribute to overall wealth inequality. Based on these findings, the paper argues that effectively addressing wealth inequality in China requires government policies targeting urban-rural and regional imbalances, while also addressing within-group inequalities according to their distinct inequality profiles."
---

@@

China's economic trajectory over the past five decades, initiated by the reform and opening-up policy launched in 1978, involved a transition from a centrally planned to a more market-oriented economy. This period yielded high average GDP growth rates averaged over 9% per year along with considerable improvements in poverty reduction, health, and education [@world_bank_world_2024]. This growth, however, manifested unevenly across time, geography, and economic sectors. Coastal provinces, for instance, often experienced higher growth rates, benefiting from earlier integration with global markets, while inland and rural areas developed more slowly, contributing to regional disparities [@liao_sixty_2016]. Sectorally, growth was initially propelled by fixed-capital investment, residential investment boom, andexport-oriented manufacturing, a model that later showed signs of creating imbalances [@world_bank_world_2024]. China's annual GDP growth rate moderated from 10.6% in 2010 to 5.2% by 2023, reflecting a slowdown from the high rates experienced earlier in the decade [@world_bank_gdp_2025].

@@

This study advances the understanding of wealth inequality in China by analyzing recent microdata from the 2017 and 2019 China Household Finance Survey (CHFS). Specifically, it evaluates whether recent wealth inequality trends align with earlier findings of declining inequality, which were predominantly based on income data [@kanbur_great_2021; @zhang_survey_2021]. The paper employs factor decomposition to identify the contributions of different asset types to overall wealth inequality. This approach is also applied separately to population subgroups defined by urban-rural and coastal-inland divisions to explore potential heterogeneity in the determinants of inequality across these dimensions. Additionally, group decomposition is used to quantify how much of total wealth inequality can be attributed to differences in average wealth between and within these subgroups. By combining recent data with these decomposition techniques, the paper provides an in-depth examination of the patterns and structural determinants underpinning wealth inequality in contemporary China. Based on these findings, this paper argues that government policies aiming to reduce China’s wealth inequality should simultaneously target urban-rural and regional imbalances, as well as address inequality within each subgroup according to their specific inequality profiles.

@@

The dataset has been utilized in numerous academic studies exploring various facets of the Chinese economy and household behavior. Prior research using CHFS data has examined topics such as income and wealth distribution [@tan_earnings_2017], rural household finance [@gan_report_2014], household responses to health shocks and insurance availability [@liu_health_2022] and many others.

@@

Regarding the causes for post-2000s decline in China’s inequality, @kanbur_great_2021 and @zhang_survey_2021 point to both the Lewis Turning Point and targeted policies. By the late 2000s, the absorption of surplus rural labor had tightened rural labor markets, narrowing the rural–urban wage gap. At the same time, the 2004 Minimum Wage Law, backed by increasingly rigorous enforcement, and the rollout of rural cooperative medical insurance in 2004, followed by rural social security schemes in 2009, together helped compress both income and wealth disparities.

## Factor decomposition of wealth inequality
This section applies the decomposition method developed by @lerman_income_1985 to identify the distinct contributions of various asset types to overall wealth inequality in China. This decomposition provides a detailed examination of how each asset component influences aggregate wealth inequality as measured by the Gini coefficient. Specifically, it separates each asset's contribution into three components: its share in total household wealth, its own Gini coefficient, and its correlation with the overall wealth rank. Formally, the overall wealth Gini coefficient can be decomposed into contributions from individual asset components as follows:

@@

where $G$ is the total wealth Gini coefficient, $S_k$ represents the share of total wealth held in asset type $k$, $G_k$ is the Gini coefficient for the distribution of asset $k$, and $R_k$ is the "Gini correlation" between asset $k$ and the rank order of total wealth. The summary statistics of each asset type are presented in @tbl-wealth-summary.

@@

Relative to the nationwide decompositions for 2017 and 2019, the regional analyses reaffirm the pervasive significance of housing, land, and financial assets in shaping wealth inequality across China, while simultaneously highlighting distinctive regional patterns. In the coastal East, the influence of housing is particularly pronounced, contributing over 80 percent (80.3%) of total inequality, approximately eight percentage points above the national average. It is even fair to say that "housing inequality is the wealth inequality" in the East region. In contrast, housing accounts for a comparatively smaller share (58.3%) in the Non-East regions. Across all three decomposition components, housing distribution in the East aligns more closely with overall wealth rankings, is more unequal, and constitutes a greater share of total household wealth.

@@

The summary statistics above indicate that both the mean and median household wealth in the East were significantly higher than those in the other regions in 2017 and 2019. Notably, in 2019, the mean wealth of households in the Western region surpassed that of the Middle region, although its median wealth remained lower. Concurrently, the Gini coefficient for the West increased sharply, becoming the highest among all regions in 2019 (0.6947). However, there is a lack of empirical literature examining the underlying causes for the rise in wealth inequality observed in the Western region during this period.

@@

The regional group decomposition indicates that, analogous to the urban-rural analysis, the between-group component accounted for over 30% of total wealth inequality in both 2017 and 2019 (36.9% and 33.8%, respectively). Between-group contributions experienced a slight decline from 2017 to 2019, mirrored by a corresponding increase in the residual component (from 32.0% to 34.5%). This shift may suggest a modest convergence in household wealth across regions. Nonetheless, it remains clear that China's wealth inequality arises significantly from both within-region disparities and regional differences, with each component consistently contributing over 30% to overall inequality.

In conclusion, the group decomposition results indicate that structural imbalances in China, primarily those between urban and rural areas and between East and Non-East regions, significantly contribute to overall wealth inequality. Each of these between-group effects individually accounts for over one-third of the total wealth inequality in 2019. Given that the residual component in each decomposition captures portions of between-group effects from the alternative grouping, the combined contribution of these structural imbalances likely surpasses one-third by some notable percentage points. Furthermore, the within-group contribution in the urban-rural decomposition is substantial, reaching 56.8% in 2019. Although the within-group contribution in the regional decomposition is relatively lower (31.5%), the combined effect of within-group inequality and residual overlap (34.5%) remains considerable. Consequently, policy efforts must not only focus on addressing structural disparities between different demographic groups but also actively mitigate inequalities within each population subgroup.

# Conclusion and policy implications
This paper analyzed wealth inequality in China using data from the 2017 and 2019 China Household Finance Survey (CHFS). The findings substantiate earlier studies focused on China's income inequality, suggesting that the country's wealth inequality has stabilized since 2010. Specifically, the national wealth Gini coefficient decreased from 0.76 in 2010 to 0.70 in 2017, and further declined to 0.67 in 2019. A comprehensive understanding of the drivers behind this moderation in inequality requires more rigorous causal analyses covering this recent period.
